---
title: "p8105_hw5_tt3022"
author: "Thomas Tang"
date: "2024-11-15"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
library
```{r}
library(ggplot2)
library(broom)
library(tidyverse)
```

Problem 2
```{r}
set.seed(1)

n <- 30            
sigma <- 5          
mu_values <- 0:6    
iterations <- 5000  
alpha <- 0.05      

perform_t_test <- function(mu) {
  data <- rnorm(n, mean = mu, sd = sigma)
  test_result <- t.test(data, mu = 0)
  tidy(test_result) %>% 
    select(estimate = estimate, p.value)
}


sim_results <- expand_grid(
  mu = mu_values,
  iter = 1:iterations
) %>%
  mutate(
    test_results = map(mu, perform_t_test)
  ) %>%
  unnest(test_results)


power_results <- sim_results %>%
  group_by(mu) %>%
  summarize(power = mean(p.value < alpha))
```

create power plot
```{r}
ggplot(power_results, aes(x = mu, y = power)) +
  geom_line() +
  geom_point() +
  labs(
    title = "Power vs Effect Size",
    x = "True Mean (μ)",
    y = "Power"
  ) +
  theme_minimal()
```

The association between effect size and power is positive and non-linear. 
As the effect size grows, the power of the test increases rapidly at first and then plateaus near 1.

```{r warning=FALSE}
average_estimates <- sim_results %>%
  group_by(mu) %>%
  summarize(
    mean_estimate = mean(estimate),
    mean_estimate_rejected = mean(estimate[p.value < alpha])
  )

ggplot(average_estimates, aes(x = mu)) +
  geom_line(aes(y = mean_estimate), color = "blue") +
  geom_line(aes(y = mean_estimate_rejected), color = "red") +
  labs(
    title = "Average Estimate of µ vs True Mean",
    x = "True Mean (μ)",
    y = "Average Estimate of μ",
    color = "Estimate Type"
  ) +
  scale_color_manual(
    values = c("blue" = "All Samples", "red" = "Rejected Samples")
  ) +
  theme_minimal()
```

The sample average of mu for tests where the null is rejected is not approximately equal to the true value of mu for smaller effect sizes.
As mu increases, this bias decreases, and the sample average of my in rejected tests becomes more representative of the true mu.


Problem 3:

```{r message=FALSE, warning=FALSE}
homicide_data <- read_csv("data/homicide-data.csv")
summary(homicide_data)
homicide_data <- homicide_data %>%
  mutate(city_state = paste(city, state, sep = ", "))

# Define unsolved homicides
unsolved_dispositions <- c("Closed without arrest", "Open/No arrest")

# Summarize data within cities
city_summary <- homicide_data %>%
  group_by(city_state) %>%
  summarize(
    total_homicides = n(),
    unsolved_homicides = sum(disposition %in% unsolved_dispositions)
  ) %>%
  arrange(desc(total_homicides))
write_csv(city_summary, "city_homicide_summary.csv")
```
There is 52,179 rows and 12 columns including identifiers, victim information, location, and disposition.

```{r}
baltimore_data <- homicide_data %>%
  filter(city_state == "Baltimore, MD")

# Total homicides and unsolved homicides in Baltimore, MD
total_homicides <- nrow(baltimore_data)
unsolved_homicides <- sum(baltimore_data$disposition %in% unsolved_dispositions)

prop_test_result <- prop.test(
  x = unsolved_homicides,  # Number of unsolved homicides
  n = total_homicides      # Total homicides
)
prop_test_summary <- tidy(prop_test_result)
print(prop_test_summary)

```
Estimated Proportion of Unsolved Homicides: 0.646
95% Confidence Interval: [0.628, 0.663]


```{r warning=FALSE}
unsolved_dispositions <- c("Closed without arrest", "Open/No arrest")

city_prop_test_results <- homicide_data %>%
  mutate(city_state = paste(city, state, sep = ", ")) %>%
  group_by(city_state) %>%
  summarize(
    total_homicides = n(),
    unsolved_homicides = sum(disposition %in% unsolved_dispositions)
  ) %>%
  mutate(
    prop_test = map2(unsolved_homicides, total_homicides, ~ prop.test(.x, .y)),
    tidy_results = map(prop_test, broom::tidy)
  ) %>%
  unnest(tidy_results) %>%
  select(city_state, estimate, conf.low, conf.high)
```
To create a plot:
```{r}
ggplot(city_prop_test_results, aes(x = reorder(city_state, estimate), y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
  coord_flip() +
  labs(
    title = "Proportion of Unsolved Homicides by City",
    x = "City",
    y = "Estimated Proportion of Unsolved Homicides"
  ) +
  theme_minimal()
```





